Convolutional equivalent layer for gravity data processing

### 2020-05-05 manuscript/abstract.tex
changes made by Val

We have developed an efficient and very fast equivalent-layer technique for gravity data processing by modifying an 
iterative method grounded on excess mass constraint that does not require the solution of linear systems. 
Taking advantage of the symmetric Block-Toeplitz Toeplitz-block (BTTB) structure of the sensitivity matrix, that raises 
when regular grids of observation points and equivalent sources (point masses) are used to set up a fictitious 
equivalent layer, we have developed an algorithm which greatly reduces the number of flops and RAM memory necessary 
to estimate a 2D mass distribution over the equivalent layer. The structure of symmetric BTTB matrix consists of 
the elements of the first column of the sensitivity matrix, which in turn can be embedded into a symmetric 
Block-Circulant Circulant-Block (BCCB) matrix. Likewise, only the first column of the BCCB matrix is needed 
to reconstruct the full sensitivity matrix completely. From the first column of BCCB matrix, its eigenvalues 
can be calculated using the 2D Fast Fourier Transform (2D FFT), which can be used to readily 
compute the matrix-vector product of the forward modeling in the fast equivalent-layer technique. As a result, 
our method is efficient to process very large datasets using either fine- or mid-grid meshes. The larger the 
dataset, the faster and more efficient our method becomes compared to the available equivalent-layer techniques. 
Tests with synthetic data demonstrate the ability of our method to satisfactorily upward-
and downward-continuing the gravity data.
Our results show very small border effects and noise amplification compared to those 
produced by the classical approach in the Fourier domain.
Besides, they show that while the running time of our method is $\approx 30.9$ seconds for 
processing $N = 1\,000\,000$ observations, the fast equivalent-layer technique spent 
$\approx 46.8$ seconds with $N = 22\,500$.
A test with field data from Caraj{\'a}s Province, Brazil, illustrates the low computational 
cost of our method to process a large data set composed of $N = 250\,000$ observations.

>>>>

We have developed an efficient and very fast equivalent-layer technique for 
gravity data processing by modifying an iterative method grounded on excess mass 
constraint that does not require the solution of linear systems. 
Taking advantage of the symmetric Block-Toeplitz Toeplitz-block (BTTB) structure 
of the sensitivity matrix, that raises when regular grids of observation points 
and equivalent sources (point masses) are used to set up a fictitious 
equivalent layer, we have developed an algorithm which greatly reduces the number o
f flops and RAM memory necessary to estimate a 2D mass distribution over the 
equivalent layer. 
The structure of symmetric BTTB matrix consists of the elements of the first column of 
the sensitivity matrix, which in turn can be embedded into a symmetric 
Block-Circulant Circulant-Block (BCCB) matrix. 
Likewise, only the first column of the BCCB matrix is needed 
to reconstruct the full sensitivity matrix completely. From the first column of 
BCCB matrix, its eigenvalues can be calculated using the 2D Fast Fourier Transform 
(2D FFT), which can be used to readily compute the matrix-vector product of the 
forward modeling in the fast equivalent-layer technique. 
As a result, our method is efficient to process very large datasets using either 
fine- or mid-grid meshes. 
The larger the dataset, the faster and more efficient our method becomes compared 
to the available equivalent-layer techniques. 
Tests with synthetic data demonstrate the ability of our method to satisfactorily upward-
and downward-continuing the gravity data.
Our results show very small border effects and noise amplification compared to those 
produced by the classical approach in the Fourier domain.
Besides, they show that while the running time of our method is $\approx 30.9$ seconds for 
processing $N = 1,000,000$ observations, the fast equivalent-layer technique spent 
$\approx 46.8$ seconds with $N = 22,500$.
A test with field data from Caraj{\'a}s Province, Brazil, illustrates the low computational 
cost of our method to process a large data set composed of $N = 250,000$ observations.

### 2020-05-05 manuscript/conclusions.tex
changes made by Val

We show that the sensitivity matrix associated with the equivalent layer technique 
for processing gravity data has a BTTB structure when the following conditions are
satisfied: (i) the observed data are disposed at a regular grid on a horizontal plane
and (ii) the equivalent sources (point masses) are placed at a constant depth, one directly 
beneath each observation point.
By exploring the BTTB structure of the sensitivity matrix, we formulate the forward modeling as a 
fast FFT convolution that requires only one column of the sensitivity matrix 
and propose an efficient approach for optimizing the computational time of an 
iterative method called fast equivalent-layer technique.
This iterative method is grounded on the excess mass constraint and does not demand the solution of 
linear systems.
Compared to the fast equivalent-layer technique, 
our approach greatly reduces the number of flops and the RAM memory necessary to estimate a 
2D mass distribution within a planar equivalent layer that fits the observed gravity data. 
For example, the number of flops is reduced by two orders 
of magnitude when processing one million of observations by using our approach. 
Traditionally, such amount of data impractically requires $7.6$ Terabytes of RAM memory to handle 
the full sensitivity matrix whereas our method takes only $61.035$ Megabytes.
This drastic reduction comes from the fact that, by using only one column of the sensitivity matrix,
our method is able to compute the gravity data produced by the whole equivalent layer with a single 
point mass via FFT.
We have successfully applied the proposed method to compute the upward/downward continuation of 
synthetic gravity data. 
Our results show that our method produces considerably smaller border effects and 
noise amplification compared to the classical Fourier approach, 
which computes the Fourier transform of the observed data.
Application to field data over the Caraj{\'a}s Province, north of Brazil, 
confirms the potential of our approach in processing an airborne gravity data set with $250\,000$ observations in approximately $0.2$ seconds. 
Our method can also be applied to process large marine gravity data sets. 
Further studies need to be carried out in order to apply our method to process 
other potential-field data.

>>>>

We show that the sensitivity matrix associated with the equivalent layer technique 
for processing gravity data has a BTTB structure when the following conditions are
satisfied: (i) the observed data are disposed at a regular grid on a horizontal plane
and (ii) the equivalent sources (point masses) are placed at a constant depth, one directly 
beneath each observation point.
By exploring the BTTB structure of the sensitivity matrix, we formulate the forward modeling as a 
FFT convolution that requires only one column of the sensitivity matrix; 
hence, the forward modeling uses, in practice, a single equivalent source,
which results in a drastic reduction of the required RAM memory. 
Next, we propose an efficient approach for optimizing the computational time of an 
iterative method called fast equivalent-layer technique.
This iterative method is grounded on the excess mass constraint and does not demand the solution of 
linear systems.
Compared to the fast equivalent-layer technique, 
our approach greatly reduces the number of flops and the RAM memory necessary to estimate a 
2D mass distribution within a planar equivalent layer that fits the observed gravity data. 
For example, the number of flops is reduced by two orders 
of magnitude when processing one million of observations by using our approach. 
Traditionally, such amount of data impractically requires $7.6$ Terabytes of RAM memory to handle 
the full sensitivity matrix whereas our method takes only $61.035$ Megabytes.
We have successfully applied the proposed method to compute the upward/downward continuations of 
synthetic gravity data. 
Our results show that our method produces considerably smaller border effects and 
noise amplification compared to the classical Fourier approach, 
which computes the Fourier transform of the observed data.
Application to field data over the Caraj{\'a}s Province, north of Brazil, 
confirms the potential of our approach in processing an airborne gravity data set with $250,000$ observations in approximately $0.2$ seconds. 
Our method can also be applied to process large marine gravity data sets. 
Further studies need to be carried out in order to apply our method to process 
other potential-field data.

### 2020-05-05 manuscript/field_data.tex
changes made by Val

We applied our method to airborne gravity data from Caraj{\'a}s, north of Brazil, which were provided by 
the Geological Survey of Brazil (CPRM). The data were collected along $131$ north-south flight lines separated 
by $3$ km and $29$ east-west tie lines separated by $12$ km.
This data set was divided in two different areas, collected in different times, having samples spacing of 
$7.65$ m and $15.21$ m along the lines, totalizing  $5\,492\,551$ observation points at a fixed height 
of $900$ m ($z_{1} = -900$ m). 
The gravity data were interpolated (Figure \ref{fig:carajas_real_data}) into a regularly spaced grid of 
$500 \times 500$ observation points ($N = 250\,000$) with a grid spacing of $716.9311$ m north-south and 
$781.7387$ m east-west.

To apply our method, we set an equivalent layer at $z_{0} = 300$ m. 
Figure \ref{fig:carajas_gz_predito_val}a shows the predicted data obtained with our method after 
$N^{it} = 50$ iterations.
The residuals (Figure \ref{fig:carajas_gz_predito_val}b), defined as the difference between the observed 
(Figure \ref{fig:carajas_real_data}) and predicted (Figure \ref{fig:carajas_gz_predito_val}a) data, show a 
very good data fit with mean close to zero ($0.0003$ mGal) and small standard deviation ($0.1160$ mGal), 
which corresponds to approximately $0.1$ \% of the maximum amplitude of the gravity data.
By using the estimated mass distribution (not shown), we performed an upward-continuation of the 
observed gravity data to a horizontal plane located $5\,000$ m above. Figure \ref{fig:up2000_carajas_500x500} 
shows a very consistent upward-continued gravity data, with a clear attenuation of the short 
wavelengths. By using our approach, the processing of the $250\,000$ observations took only 
$0.216$ seconds.

>>>>

We applied our method to airborne gravity data from Caraj{\'a}s, north of Brazil, which were provided by 
the Geological Survey of Brazil (CPRM). The data were collected along $131$ north-south flight lines separated 
by $3$ km and $29$ east-west tie lines separated by $12$ km.
This data set was divided in two different areas, collected in different times, having samples spacing of 
$7.65$ m and $15.21$ m along the lines, totalizing  $5,492,551$ observation points at a fixed height 
of $900$ m ($z_{1} = -900$ m). 
The gravity data were interpolated (Figure \ref{fig:carajas_real_data}) into a regularly spaced grid of 
$500 \times 500$ observation points ($N = 250,000$) with a grid spacing of 
$\approx 717$ m north-south and $\approx 782$ m east-west.

To apply our method, we set an equivalent layer at $z_{0} = 300$ m. 
Figure \ref{fig:carajas_gz_predito_val}a shows the predicted data obtained with our method after 
$N^{it} = 50$ iterations.
The residuals (Figure \ref{fig:carajas_gz_predito_val}b), defined as the difference between the observed 
(Figure \ref{fig:carajas_real_data}) and predicted (Figure \ref{fig:carajas_gz_predito_val}a) data, show a 
very good data fit with mean close to zero ($0.0003$ mGal) and small standard deviation ($0.1160$ mGal), 
which corresponds to approximately $0.1$ \% of the maximum amplitude of the gravity data.
By using the estimated mass distribution (not shown), we performed an upward-continuation of the 
observed gravity data to a horizontal plane located $5,000$ m above. Figure \ref{fig:up2000_carajas_500x500} 
shows a very consistent upward-continued gravity data, with a clear attenuation of the short 
wavelengths. By using our approach, the processing of the $250,000$ observations took only 
$0.216$ seconds.

### 2020-05-05 manuscript/figures.tex
changes made by Val

\plot{Figure2}{width=\textwidth}{
	{Comparison between the number of flops (equations \ref{eq:float_fast_eqlayer} and \ref{eq:float_bccb}) 
	associated with the fast equivalent-layer technique \citep{siqueira-etal2017} and our method, 
	for $N$ varying from $5\,000$ to $1\,000\,000$. All values are computed with $N^{it} = 50$ 
	iterations and $\kappa = 5$.}
	\label{fig:float}
}

@@

\plot{synthetic_data_med}{width=13.5cm}{
	{Application to synthetic data. 
	Noise-corrupted gravity data (in color map) produced by three synthetic sources: 
	a sphere with density contrast $-1.25\, \mathrm{g/cm^3}$ and 
	two rectangular prisms with density contrasts $1.00\, \mathrm{g/cm^3}$ (upper-left body) and 
	$1.30\, \mathrm{g/cm^3}$ (upper-right body). 
	The synthetic data are contaminated with a pseudorandom Gaussian noise having null mean 
	and standard deviation equal to $0.1 \, \mathrm{mGal}$.
	The black lines represent the horizontal projection of the sources.}
	\label{fig:synthetic_data}
}

\plot{deltas_gz_med}{width=6cm}{
	{Application to synthetic data. 
	(a) Residuals between the simulated data (Figure \ref{fig:synthetic_data}) and 
	predicted data produced by the fast equivalent-layer technique \citep{siqueira-etal2017}. 
	(b) Residuals between the simulated data (Figure \ref{fig:synthetic_data}) and 
	predicted data produced by our method. 
	The mean ($-4.493 \times 10^{-5}$ mGal) and standard deviation ($0.093$ mGal) for residuals
	shown in a and b are exactly the same.
	(c) Difference between the predicted data produced by the fast equivalent-layer technique 
	and our method.
	The computation times spent by the fast equivalent-layer technique and 
	our method were $10.416$ and $0.177$ seconds, respectively.}
	\label{fig:comparison_datafits}
}

\plot{upward_fourier_med}{width=13.5cm}{
	{Application to synthetic data.
	(a) Noise-free gravity data produced by the synthetic sources at $z = -300$ m.
	(b) Residuals between the data shown in a and the upward-continued 
	data obtained with our method (not shown), with 
	mean $0.003$ mGal and standard deviation $0.034$ mGal. 
	(c) Residuals between the data shown in a and the upward-continued 
	data obtained with the fast equivalent-layer technique (not shown), with 
	mean $0.003$ mGal and standard deviation $0.034$ mGal.
	(d) Residuals between the data shown in a and the upward-continued 
	data obtained with the classical Fourier approach (not shown), with 
	mean $-0.030$ mGal and standard deviation $0.262$ mGal.
	Values shown in b, c and d have the same scale. 	 
	The computation times spent by the fast equivalent-layer technique and 
	our method were $8.697$ and $0.005$ seconds, respectively.}
	\label{fig:upward_comparison}
}

\plot{downward_fourier_med}{width=13.5cm}{
	{Application to synthetic data. 
	(a) Noise-free gravity data produced by the synthetic sources at $z = -50$ m.
	(b) Residuals between the data shown in a and the downward-continued 
	data obtained with our method (not shown), with 
	mean $-0.001$ mGal and standard deviation $0.038$ mGal. 
	(c) Residuals between the data shown in a and the downward-continued 
	data obtained with the fast equivalent-layer technique (not shown), with 
	mean $-0.001$ mGal and standard deviation $0.038$ mGal.
	(d) Residuals between the data shown in a and the downward-continued 
	data obtained with the classical Fourier approach (not shown), with 
	mean $-0.030$ mGal and standard deviation $0.262$ mGal.
	Values shown in b, c and d have the same scale.	 
	The computation times spent by the fast equivalent-layer technique and 
	our method were $8.795$ and $0.004$ seconds, respectively.}
	\label{fig:downward_comparison}
}

@@

\plot{Figure14}{width=13.5cm}{
	{Application to field data over the Caraj{\'a}s Province, Brazil. 
	The upward-continued gravity data obtained with our method $5\,000$ m above the 
	observed data (Figure \ref{fig:carajas_real_data}).
	The total computation time for processing of the $250,000$ observations was $0.216$ seconds.}
	\label{fig:up2000_carajas_500x500}
}

>>>>

\plot{Figure2}{width=\textwidth}{
	{Comparison between the number of flops (equations \ref{eq:float_fast_eqlayer} and \ref{eq:float_bccb}) 
	associated with the fast equivalent-layer technique \citep{siqueira-etal2017} and our method, 
	for $N$ varying from $5,000$ to $1,000,000$. All values are computed with $N^{it} = 50$ 
	iterations and $\kappa = 5$.}
	\label{fig:float}
}

@@

\plot{synthetic_data_med}{width=13.5cm}{
	{Application to synthetic data. 
	Noise-corrupted gravity data (in color map) produced by three synthetic sources: 
	a sphere with density contrast $-1.25\, \mathrm{g/cm^3}$ and 
	two rectangular prisms with density contrasts $1.00\, \mathrm{g/cm^3}$ (upper-left body) and 
	$1.30\, \mathrm{g/cm^3}$ (upper-right body). 
	The black lines represent the horizontal projection of the sources.}
	\label{fig:synthetic_data}
}

\plot{deltas_gz_med}{width=6cm}{
	{Application to synthetic data. 
	Residuals between the simulated data (Figure \ref{fig:synthetic_data}) and 
	predicted data produced by: (a) the fast equivalent-layer technique \citep{siqueira-etal2017} and (b) our method. 
	The mean ($-4.493 \times 10^{-5}$ mGal) and standard deviation ($0.093$ mGal) for residuals
	shown in a and b are exactly the same.
	(c) Difference between the predicted data produced by the fast equivalent-layer technique 
	and our method.
	The computation times spent by the fast equivalent-layer technique and 
	our method were $10.416$ and $0.177$ seconds, respectively.}
	\label{fig:comparison_datafits}
}

\plot{upward_fourier_med}{width=13.5cm}{
	{Application to synthetic data.
	(a) Noise-free gravity data produced by the synthetic sources at $z = -300$ m.
	Residuals between the data shown in a and the upward-continued 
	data obtained by: (b) our method (not shown), with 
	mean $0.003$ mGal and standard deviation $0.034$ mGal, 
	(c) the fast equivalent-layer technique (not shown), with 
	mean $0.003$ mGal and standard deviation $0.034$ mGal and 
	(d) the classical Fourier approach (not shown), with 
	mean $-0.030$ mGal and standard deviation $0.262$ mGal.
	Values shown in b, c and d have the same scale. 	 
	The computation times spent by the fast equivalent-layer technique and 
	our method were $8.697$ and $0.005$ seconds, respectively.}
	\label{fig:upward_comparison}
}

\plot{downward_fourier_med}{width=13.5cm}{
	{Application to synthetic data. 
	(a) Noise-free gravity data produced by the synthetic sources at $z = -50$ m.
	Residuals between the data shown in a and the downward-continued 
	data obtained by: (b) our method (not shown), with 
	mean $-0.001$ mGal and standard deviation $0.038$ mGal, 
	(c) the fast equivalent-layer technique (not shown), with 
	mean $-0.001$ mGal and standard deviation $0.038$ mGal and 
	(d) the classical Fourier approach (not shown), with 
	mean $-0.030$ mGal and standard deviation $0.262$ mGal.
	Values shown in b, c and d have the same scale.	 
	The computation times spent by the fast equivalent-layer technique and 
	our method were $8.795$ and $0.004$ seconds, respectively.}
	\label{fig:downward_comparison}
}

@@

\plot{Figure14}{width=13.5cm}{
	{Application to field data over the Caraj{\'a}s Province, Brazil. 
	The upward-continued gravity data obtained with our method $5,000$ m above the 
	observed data (Figure \ref{fig:carajas_real_data}).
	The total computation time for processing of the $250,000$ observations was $0.216$ seconds.}
	\label{fig:up2000_carajas_500x500}
}

### 2020-05-05 manuscript/synthetic_data.tex
changes made by Val

We have simulated three sources whose horizontal projections are shown in 
Figure \ref{fig:synthetic_data} as black lines. 
These sources are a sphere with density contrast  
$-1.25\, \mathrm{g/cm^3}$ and two rectangular prisms with density contrasts 
$1.00\, \mathrm{g/cm^3}$ (upper-left prism) 
and $1.30\, \mathrm{g/cm^3}$ (upper-right prism). 
Figure \ref{fig:synthetic_data} shows the gravity disturbance (vertical component of 
gravitational attraction) produced by these sources. 
The synthetic data are contaminated with additive pseudorandom Gaussian 
noise with zero mean and standard deviation of $0.1 \, \mathrm{mGal}$.
The data are computed at $N = 10\,000$ observation points that are regularly spaced on a 
$100 \times 100$ grid, at $z_{1} = -100$ m. 
We have set a grid of equivalent sources, each one directly beneath each 
observation point, at $z_{0} = 300$ m.

Figures \ref{fig:comparison_datafits}a and \ref{fig:comparison_datafits}b show the data fits 
obtained, respectively, by the fast equivalent-layer technique \citep{siqueira-etal2017} 
and by our method. They represent the differences between the simulated data (Figure \ref{fig:synthetic_data}) and the predicted data produced by both methods 
(not shown) after $N^{it} = 40$ iterations. 
As we can see, both methods produce virtually the same results.
This excellent agreement is confirmed by Figure \ref{fig:comparison_datafits}c, which shows 
the differences between the predicted data produced by both methods.

We performed the upward and downward continuations of the simulated gravity data 
(Figure \ref{fig:synthetic_data}) by using the fast equivalent-layer technique
\citep{siqueira-etal2017}, our method and also the classical approach in the 
Fourier domain. This approach consists in performing the upward/downward continuation 
by directly computing the Fourier transform of the gravity data 
\citep[e.g., ][ p. 317]{blakely1996}.
Figure \ref{fig:upward_comparison} shows the upward-continued gravity data 
obtained by the three methods.
As we can see, the residuals between the true data at $z = -300$ m 
(Figure \ref{fig:upward_comparison}a) and the upward-continued data obtained by 
using our method (Figure \ref{fig:upward_comparison}b) and 
the fast equivalent-layer technique (Figure \ref{fig:upward_comparison}c) are very similar 
to each other.
Notice that the absolute values of the residuals produced by the classical Fourier 
approach (Figure \ref{fig:upward_comparison}d) are $\approx 10$ times greater than
those produced by our method and the fast equivalent-layer technique 
(Figures \ref{fig:upward_comparison}b and \ref{fig:upward_comparison}c), 
with maximum values concentrated at the border of the simulated area.
We opted for showing the results produced by all methods without removing the border 
effects in order to properly compare them to each other.
We can also notice that, differently from the results produced by the other methods, 
that obtained with the classical Fourier approach exhibits a slight noise amplification.

Figure \ref{fig:downward_comparison} shows the results obtained by using all 
methods to compute downward continuation of gravity data.
In this case, the maximum absolute values of the residuals produced by the 
classical Fourier approach (Figure \ref{fig:downward_comparison}d) are 
$\approx 20$ times greater than those produced by our method and the fast equivalent-layer technique (Figures \ref{fig:downward_comparison}b and \ref{fig:downward_comparison}c).
It is also important to stress that the result produced by the classical Fourier approach 
exhibits a considerable noise amplification when compared to the other methods.
This noise amplification is a well-known problem of the downward continuation 
produced by the classical Fourier approach \citep[e.g., ][ p. 320]{blakely1996}.

Another important aspect to be pointed out about the results we obtained with 
synthetic data (Figures \ref{fig:upward_comparison} and \ref{fig:downward_comparison})
is the computational times spent by our method and the fast equivalent-layer technique. 
The total computational time required by our method to estimate the physical property 
distribution on the equivalent layer and perform the upward/downward continuation 
is about two orders of magnitude lower than that spent by the fast equivalent-layer technique.
This significant reduction in computational time was obtained by using a data set 
composed of $N = 10\,000$ observation points. Considerably better results can be 
obtained with larger data sets.

Finally, we did not compare the total computational times spent by our method and the classical 
Fourier approach, but we can affirm that the second is smaller. 
While the classical approach needs basically one DFT/IDFT of the data, 
our method requires one DFT/IDFT per iteration. 
On the other hand, the considerably smaller noise amplification and 
border effect are the main advantages of our method over the classical Fourier 
approach, especially in the downward continuation.

>>>>

We have simulated three sources whose horizontal projections are shown in 
Figure \ref{fig:synthetic_data} as black lines. 
These sources are a sphere with density contrast  
$-1.25\, \mathrm{g/cm^3}$ and two rectangular prisms with density contrasts 
$1.00\, \mathrm{g/cm^3}$ (upper-left prism) 
and $1.30\, \mathrm{g/cm^3}$ (upper-right prism). 
Figure \ref{fig:synthetic_data} shows the gravity disturbance (vertical component of 
gravitational attraction) produced by these sources. 
The synthetic data are contaminated with additive pseudorandom Gaussian 
noise with zero mean and standard deviation of $0.1 \, \mathrm{mGal}$.
The data are computed at $N = 10,000$ observation points that are regularly spaced on a 
$100 \times 100$ grid, at $z_{1} = -100$ m. 
We have set a grid of equivalent sources, each one directly beneath each 
observation point, at $z_{0} = 300$ m.

Figure \ref{fig:comparison_datafits}a and \ref{fig:comparison_datafits}b show the data fits 
obtained, respectively, by the fast equivalent-layer technique \citep{siqueira-etal2017} 
and by our method. They represent the differences between the simulated data (Figure \ref{fig:synthetic_data}) and the predicted data produced by both methods 
(not shown) after $N^{it} = 40$ iterations. 
As we can see, both methods produce virtually the same results.
This excellent agreement is confirmed by Figure \ref{fig:comparison_datafits}c, which shows 
the differences between the predicted data produced by both methods.

We performed the upward and downward continuations of the simulated gravity data 
(Figure \ref{fig:synthetic_data}) by using the fast equivalent-layer technique
\citep{siqueira-etal2017}, our method and also the classical approach in the 
Fourier domain. This approach consists in performing the upward/downward continuation 
by directly computing the Fourier transform of the gravity data 
\citep[e.g., ][ p. 317]{blakely1996}.
Figure \ref{fig:upward_comparison} shows the upward-continued gravity data 
obtained by the three methods.
As we can see, the residuals between the true data at $z = -300$ m 
(Figure \ref{fig:upward_comparison}a) and the upward-continued data obtained by 
using our method (Figure \ref{fig:upward_comparison}b) and 
the fast equivalent-layer technique (Figure \ref{fig:upward_comparison}c) are very similar 
to each other.
Notice that the absolute values of the residuals produced by the classical Fourier 
approach (Figure \ref{fig:upward_comparison}d) are $\approx 10$ times greater than
those produced by our method and the fast equivalent-layer technique 
(Figure \ref{fig:upward_comparison}b and \ref{fig:upward_comparison}c), 
with maximum values concentrated at the border of the simulated area.
Differently from the results yield by our method (Figure \ref{fig:upward_comparison}b) 
and the fast equivalent-layer technique (Figure \ref{fig:upward_comparison}c), 
that obtained with the classical Fourier approach exhibits a slight noise 
amplification (Figure \ref{fig:upward_comparison}d).

Figure \ref{fig:downward_comparison} shows the results obtained by using all 
methods to compute downward continuation of gravity data.
In this case, the maximum absolute values of the residuals produced by the 
classical Fourier approach (Figure \ref{fig:downward_comparison}d) are 
$\approx 20$ times greater than those produced by our method and the fast equivalent-layer technique (Figure \ref{fig:downward_comparison}b and \ref{fig:downward_comparison}c).
This noise amplification is a well-known problem of the downward continuation 
produced by the classical Fourier approach \citep[e.g., ][ p. 320]{blakely1996}.

We opted for showing all the results (Figures \ref{fig:upward_comparison} and 
\ref{fig:downward_comparison}) produced by all methods without removing the border 
effects in order to properly compare them to each other.
We also stress that no padding function to expand the data was used in applying 
our method, the fast equivalent-layer technique or the Fourier approach.
Another important aspect to be pointed out about these results 
is the computational times spent by our method and the fast equivalent-layer technique. 
The total computational time required by our method to estimate the physical-property 
distribution on the equivalent layer and to perform the upward/downward continuations 
is about two orders of magnitude lower than that spent by the fast equivalent-layer technique.
This significant reduction in computational time was obtained by using a data set 
composed of $N = 10,000$ observation points. Considerably better results can be 
obtained with larger data sets.

Finally, we did not compare the total computational times spent by our method 
and the classical Fourier approach, but we can affirm that the second is smaller. 
Because  the Fourier approach requires only one DFT/IDFT of the data, whereas 
our method requires one DFT/IDFT per iteration, it is computationally faster 
than our method.
However, the considerably smaller noise amplification and practically nonexistent 
border effect are the main advantages of our method over the classical Fourier 
approach, especially in the downward continuation.
